Catchy Title

### 2011-10-11 proposal/paper.tex
Kai's modification

\section{Introduction}
\label{sec:intro}
Timeseries data are prevalent in large-scale computing centers. Systems often capture sampled metrics of performance, utilization, and even sensor data like temperature. These streams are used for monitoring, placement, optimization, and more: for example, task assignment algorithms use computer utilization to determine where to place jobs.

@@

The scope of potential applications is considerable, and there is prior art in models for these operations on timeseries data (Sec. \ref{sec:survey}). We plan to focus our efforts on applying and tuning existing algorithms from the literature on the different type of timeseries data we have: rather than target  relatively small sensor datasets as many have considered, we seek to apply these techniques to lengthy, bursty utilization data. This focus motivates our emphasis on evaluating multiple models, preprocessing, and examining scalability (Sec. \ref{sec:method}).

@@

SPIRIT \cite{papadimitriou2005streaming} is an algorithm for discovering patterns in multiple streams of time-series data. This algorithm effectively tries to find correlations between the different data streams by performing principal component analysis on the the data covariance matrix. These principal components are modified and updated as each data sample is received at each time step while forgetting the samples. An interesting aspect of this algorithm is, anomaly detection which is done by detecting the changes in the number of principal components required to capture most of the energy in the signal.
In \cite{roweis1999unifying} the authors describe a unifying model for Linear Gaussian Models which includes PCA, Kalman Filters, Gaussian Mixture Models, ICA etc. They describe a general linear state transition model along with a linear observation model corrupted by independent Gaussian noise. They describe general algorithms to do inference and also learn the parameters of the model.

@@

Zhu and Shasha \cite{Zhu2003} address the problem of detecting bursts in time series when their time length are unknown. The core approach is to convert time series data to wavelets, and maintain them  for a given sliding window of time in a shifted wavelet tree. This data structure allows constructing a pattern of a flexible size within that window from wavelets of different sizes. In turn, it enables analyzing the time series data at different time scales by extracting patterns of different sizes. An online algorithm for maintaining and searching the data structure gives a constant time for updating it, while a batch algorithm trades it off for constant search times. This work focuses on efficiently detecting bursts in one time series. On the other hand, we would need to extend their technique and apply it to a number of time series. We may need to examine bursts in a group of time series, not only those in individual series.

Sakurai et al. \cite{Sakurai2011} introduce an effective way of finding trends in large time series data. In particular, they propose a method for selecting an appropriate window size in time that analysts would find useful. The idea is based on first using independent component analysis (ICA), and then selecting the window size that maximizes the entropy of the weight values in the mixing matrix. By using this criterion, the technique effectively extracts repetitive patterns at different time scales from various data sets, such as Ondemand TV access records and web-click traces for a search engine. This work is complementary to the burst detection work by Zhu and Shasha, as the former suggests the best window size that exhibits repetitive patterns while the latter allows examining different window sizes efficiently. Although the paper focuses on web-click records, the general technique would also be applicable to our data center logs.

@@

PAL is an anomaly localization tool to detect faulty components in distributed applications. The key idea is firstly to monitor time-series metrics of each individual component of the distributed system and discover the change points (potentially some faulty behavior happen), and then correlated different change points based on time stamp to locate the real faulty component. Their algorithm has four steps: firstly, raw time series data needs to be smoothed. Kalman filter failed to capture the critical change points so they used 5-length moving average filter. The second step is to use CUSUM to find change points, and calculate the “separation level” of each change point (which measures the different of mean value of data points before the change point and after it). The outlier change points (whose separation level are outliers) is the critical change points.

Another console logs mining paper \cite{xu2009} extracts specific features from unstructured logs of Hadoop system \cite{Hadoop} (A distributed system for data intensive computing) and use them to detect anomalies in the system. The tool is mainly used to detect “correctness” problems. The author combines logs and source to extract three different signals: “message types”, “state variables”, and “message identifiers”. Based on these signals, they create two matrices: one matrix has “state variables” as the column, and time as the row, the other matrix uses “message types” as the column, and “message identifiers” as the row. Finally, PCA is used for anomaly detection. They perform PCA on the two matrices to get a reduced set of dimensions (eigenvectors) ranked by how much variance in the data each reduced dimension explains. Anomalous rows are identified by flagging those rows whose data is largely unexplained by the first K eigenvectors, which describes 95\% of the overall observed variance. 
After finding the anomalous rows, human need to manually interpret and further digging the root causes. In \cite{xu2010}, they further applied their algorithms into some of Google’s production logs. Since the data is large which makes it unrealistic to perform PCA, they used frequent mining techniques to segment the long-lasting sequence into sessions.

@@

\subsection{Data Collection}
\label{sec:method:data}
The data sets is mainly collected from two data centers: OpenCloud and DCO (?). OpenCloud is a CMU-based academic cluster consisting of 64 compute nodes, each with dual quad-core 2.83GHz Intel Xeon processors, 16GB memory and a 10GigE NIC, and Arista 10 GigE switches. OpenCluster mainly runs distributed services like Hadoop file systems and map reduce systems. We are collecting OS-level metrics and Hadoop system logs per 5 seconds, and have data over several months.

@@

In addition to learning a model and using it to predict future patterns, we plan to do an analysis on existing data to observe important trends, such as bursts and repetitive patterns. Our main challenge in this problem is to deal with large data sets efficiently and flexibly. Our proposed direction is to build on the burst detection technique by Zhu et al. \cite{Zhu2003} and learn trends at different time scales and possibly with varied granularity of data aggregation across data center nodes.

@@

\subsection{Scalability}
\label{sec:method:scale}
As a second phase of the project, we plan to choose the model that works best for our data on the basis of how accurately it can provide short-term predictions. As observed above, the primary challenge for many of these algorithms is efficiency and scale. Though the precise steps we will take to improve algorithm scalability at this point are ill-defined, we do plan to investigate scalability and consider implementation on Hadoop (perhaps with simple sharding of the data).

\subsection{Anomaly Detection}
\label{sec:method:anomaly}
After validating the right model for our timeseries data from the data center, we plan to apply the model and combine it with anomaly detection algorithm. The first stage is to manually inject failures to tune the model to find out the anomalies. Later on, we can apply the algorithms on historical data from OpenCloud data center to test whether the algorithm can find anomalies in the system.

>>>>

\section{Introduction}
\label{sec:intro}
Timeseries data are prevalent in large-scale computing centers. Systems often capture sampled metrics of performance, utilization, and even sensor data like temperature. System logs can be also extracted to event signal associated with time stamp. These streams are used for monitoring, placement, optimization, and more: for example, task assignment algorithms use computer utilization to determine where to place jobs.

@@

The scope of potential applications is considerable, and there is prior art in models for these operations on timeseries data (Sec. \ref{sec:survey}). We plan to focus our efforts on applying and tuning existing algorithms from the literature on the different type of timeseries data we have: rather than target  relatively small sensor datasets as many have considered, we seek to apply these techniques to lengthy, bursty utilization data which are collected from real production clusters. This focus motivates our emphasis on evaluating multiple models, preprocessing, and examining scalability (Sec. \ref{sec:method}).

@@

SPIRIT \cite{papadimitriou2005streaming} is an algorithm for discovering patterns in multiple streams of timeseries data. This algorithm effectively tries to find correlations between the different data streams by performing principal component analysis on the the data covariance matrix. These principal components are modified and updated as each data sample is received at each time step while forgetting the samples. An interesting aspect of this algorithm is, anomaly detection which is done by detecting the changes in the number of principal components required to capture most of the energy in the signal.
In \cite{roweis1999unifying} the authors describe a unifying model for Linear Gaussian Models which includes PCA, Kalman Filters, Gaussian Mixture Models, ICA etc. They describe a general linear state transition model along with a linear observation model corrupted by independent Gaussian noise. They describe general algorithms to do inference and also learn the parameters of the model.

@@

Zhu and Shasha \cite{Zhu2003} address the problem of detecting bursts in timeseries when their time length are unknown. The core approach is to convert time series data to wavelets, and maintain them  for a given sliding window of time in a shifted wavelet tree. This data structure allows constructing a pattern of a flexible size within that window from wavelets of different sizes. In turn, it enables analyzing the timeseries data at different time scales by extracting patterns of different sizes. An online algorithm for maintaining and searching the data structure gives a constant time for updating it, while a batch algorithm trades it off for constant search times. This work focuses on efficiently detecting bursts in one timeseries. On the other hand, we would need to extend their technique and apply it to a number of timeseries. We may need to examine bursts in a group of timeseries, not only those in individual series.

Sakurai et al. \cite{Sakurai2011} introduce an effective way of finding trends in large timeseries data. In particular, they propose a method for selecting an appropriate window size in time that analysts would find useful. The idea is based on first using independent component analysis (ICA), and then selecting the window size that maximizes the entropy of the weight values in the mixing matrix. By using this criterion, the technique effectively extracts repetitive patterns at different time scales from various data sets, such as Ondemand TV access records and web-click traces for a search engine. This work is complementary to the burst detection work by Zhu and Shasha, as the former suggests the best window size that exhibits repetitive patterns while the latter allows examining different window sizes efficiently. Although the paper focuses on web-click records, the general technique would also be applicable to our data center logs.

@@

PAL is an anomaly localization tool to detect faulty components in distributed applications. The key idea is firstly to monitor timeseries metrics of each individual component of the distributed system and discover the change points (potentially some faulty behavior happen), and then correlated different change points based on time stamp to locate the real faulty component. Their algorithm has four steps: firstly, raw time series data needs to be smoothed. Kalman filter failed to capture the critical change points so they used 5-length moving average filter. The second step is to use CUSUM to find change points, and calculate the “separation level” of each change point (which measures the different of mean value of data points before the change point and after it). The outlier change points (whose separation level are outliers) is the critical change points.

Another console logs mining paper \cite{xu2009} extracts specific features from unstructured logs of Hadoop system \cite{Hadoop} (A distributed system for data intensive computing) and use them to detect anomalies in the system. The tool is mainly used to detect correctness problems. The author combines logs and source to extract three different signals: message types, state variables, and message identifiers. Based on these signals, they create two matrices: one matrix has state variables as the column, and time as the row, the other matrix uses message types as the column, and message identifiers as the row. Finally, PCA is used for anomaly detection. They perform PCA on the two matrices to get a reduced set of dimensions (eigenvectors) ranked by how much variance in the data each reduced dimension explains. Anomalous rows are identified by flagging those rows whose data is largely unexplained by the first K eigenvectors, which describes 95\% of the overall observed variance. 
After finding the anomalous rows, anomalies need to be manually interpreted to dig out the root causes. In \cite{xu2010}, they further applied their algorithms into some of Google's production logs. Since the data is large which makes it unrealistic to perform PCA, they used frequent mining techniques to segment the long-lasting sequence into sessions.

@@

\subsection{Data Collection}
\label{sec:method:data}
The data sets is mainly collected from one data center: OpenCloud. OpenCloud is a CMU-based academic cluster consisting of 64 compute nodes, each with dual quad-core 2.83GHz Intel Xeon processors, 16GB memory and a 10GigE NIC, and Arista 10 GigE switches. OpenCluster mainly runs distributed services like Hadoop file systems and map reduce systems. We are collecting OS-level metrics and Hadoop system logs per 5 seconds, and have data over several months.

@@

In addition to learning a model and using it to predict future patterns, we plan to do an analysis on existing data to observe important trends, such as bursts and repetitive patterns. Our main challenge in this problem is to deal with large data sets efficiently and flexibly. Our proposed direction is to build on the burst detection technique by Zhu et al. \cite{Zhu2003} and learn trends at different time scales and possibly with varied granularity of data aggregation across data center nodes.
Preprocessing

@@

\subsection{Scalability}
\label{sec:method:scale}
In order to handle multiple streams from large scale data centers, it is necessary to parallelize our analysis algorithm. We plan to use Hadoop as the computing framework to implement the parallel version of our algorithm.

\subsection{Anomaly Detection}
\label{sec:method:anomaly}
After validating the right model for our timeseries data from the data center, we plan to apply the model and combine it with anomaly detection algorithm. The first stage is to manually inject failures to tune the model to increase the accuracy. Later on, we can apply the algorithms on historical data from OpenCloud data center to test whether the algorithm can find real anomalies in the system.
